\documentclass[12pt]{article}
\usepackage{graphicx} % Required for inserting images

\title{Problem Set 5}
\author{Julie Dawkins}
\date{March 2024}

\begin{document}

\maketitle

\section*{Problem 3 - Webscraping}
This data lists 150 university towns. I am interested in studying homelessness in university towns, exploring links between enrollment, rent prices, and homelessness counts. This data is incomplete in its usefulness -- I need more data on the cities such as total population, number of students enrolled at the university, et cetera -- but it is a good starting point listing schools and their cities to help me find a more robust data set. 

I did not use any online tutorials to help me scrape this but did use my own notes from class (specifically, the test.r file in which we scraped from Wikipedia). 
\section*{Problem 4 - API}
For this problem, I signed up for a Housing and Urban Development API in order to access Fair Market Rent (FMR) data for counties and metro-areas. FMR is defined as the 40th percentile in rent for a 1-to-4 bedroom unit in that county or metropolitan area. 

I thought this data was very interesting and could be a very helpful way to quickly gather both "treated" data and controls for many regional areas. Examining the FMR for various regions yielded expected results -- e.g., the Tulsa and Oklahoma metro areas are more expensive than anywhere else in Oklahoma. 

To gather this data, I needed to use the packages "httr" and "jsonlite". 
\end{document}
